An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-09-09 CMPSpaper_writing/CMPSpaper_writing.Rmd
added codes for reproducibility; modified the Results section

@@

@@

```{r hamby252-data, eval=TRUE, echo=FALSE}
# for hamby 252
N <- 3
CMPS_hamby252_results <- list()
CMPS_hamby252_results$span1 <- list(0.75, 0.25, 0.25)
CMPS_hamby252_results$signame <-
  list("sigs75", "sigs25", "sigs25_1062")
CMPS_hamby252_results$npeaks.set <-
  list(c(5, 3, 1),
       c(5, 3, 1),
       c(10, 6, 2))
CMPS_hamby252_results$seg_length <- list(50, 50, 50)
CMPS_hamby252_results$Tx <- list(25, 25, 25)
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby252_results$seg_length[[i]],
       ", Tx=",
      CMPS_hamby252_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

for(i in 1:N){
  CMPS_hamby252_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/", 
                        CMPS_hamby252_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

com.title252 <- expression(paste(
  "Hamby 252 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

# generate plots
for (i in 1:3) {
  hamby252.cmps <- CMPS_hamby252_results$cmps.table[[i]]

  hamby252.plot.list <- list()
  titlee <- CMPS_hamby252_results$titlee[[i]]

  hamby252.plot.list[[1]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      # title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    font("x.text", size = 6) +
    scale_fill_manual(values=c("darkorange", "darkgrey"))

  hamby252.plot.list[[2]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      # title = expression(paste(
      #   "Hamby252 - ", bar(CMPS)[max], " Distribution"
      # )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    scale_fill_manual(values=c("darkorange", "darkgrey"))

  plot <- ggarrange(plotlist = hamby252.plot.list,
                    nrow = 1,
                    ncol = 2,
                    common.legend = TRUE, legend = "bottom")
  plot <- annotate_figure(plot, 
                          top = text_grob(com.title252))
  CMPS_hamby252_results$plot[[i]] <- plot
}

@@

```{r hamby44-data, eval=TRUE, echo=FALSE}
N <- 3
CMPS_hamby44_results <- list()
CMPS_hamby44_results$span1 <- list(0.25, 0.25, 0.75)
CMPS_hamby44_results$signame <-
  list("sigs25_531", "sigs25_1061", "sigs75_531")
CMPS_hamby44_results$npeaks.set <-
  list(c(5, 3, 1),
       c(10, 6, 1),
       c(5, 3, 1))
CMPS_hamby44_results$seg_length <- as.list(rep(61, 3))
CMPS_hamby44_results$Tx <- as.list(rep(30, 3))
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby44_results$seg_length[[i]],
       ", Tx=",
      CMPS_hamby44_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()

for(i in 1:N){
  CMPS_hamby44_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/", 
                        CMPS_hamby44_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

com.title44 <- expression(paste(
  "Hamby 44 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

# generate plots
for (i in 1:3) {
  hamby44.cmps <- CMPS_hamby44_results$cmps.table[[i]]

  hamby44.plot.list <- list()
  titlee <- CMPS_hamby44_results$titlee[[i]]

  hamby44.plot.list[[1]] <- hamby44.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      # title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    font("x.text", size = 6)+
    scale_fill_manual(values=c("darkorange", "darkgrey"))

  hamby44.plot.list[[2]] <- hamby44.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      # title = expression(paste(
      #   "Hamby252 - ", bar(CMPS)[max], " Distribution"
      # )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    scale_fill_manual(values=c("darkorange", "darkgrey"))

  plot <- ggarrange(plotlist = hamby44.plot.list,
                    nrow = 1,
                    ncol = 2,
                    common.legend = TRUE, legend = "bottom")
  plot <- annotate_figure(plot, 
                          top = text_grob(com.title44))
  CMPS_hamby44_results$plot[[i]] <- plot
}
```

@@

The sum of squares ratios are 20.85 and 29.32 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. \hh{This indicates, that even though scores from $\mathrm{CMPS_{max}}$ for known-match comparisons are larger than scores from the averaged version of $\mathrm{\overline{CMPS}_{max}}$, these scores achieve a better separation between the two groups of comparisons.}

```{r, echo=FALSE, eval=FALSE}
with(CMPS_hamby252_results$cmps.table[[2]], {
  tibble(cmps.max.m, cmps.maxbar.m) %>% 
    apply(2, compute_var_ratio_anova, label = type_truth, MS = FALSE)
})
```

<!-- this is figure 2 -->
```{r result1_252, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 252; removed outliers; seg\\_length = 50, Tx = 25, npeaks.set = c(5,3,1) ", fig.height=3}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
# knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
CMPS_hamby252_results$plot[[2]]
```

@@

For Hamby set 44 we not only generated comparable results but also used different sets of parameters to examine their potential effects.

\autoref{fig:result1_44} shows the Hamby 44 results that are comparable with Hamby 252 results. The parameters used in `extract_feature_cmps` are:

@@

Since the Hamby set 44 scans have a different scanning resolution, we need to adjust the value of `seg_length` and `Tx` accordingly. One unit in Hamby 44 represents 1.29 \textmu m. Thus, we make `seg_length = 61` and `Tx = 30` to ensure that each basis segment has length 78.69 \textmu m, and the tolerance zone is $\pm 38.7$ \textmu m. Again, we are able to see a clear separation between the KM comparisons and the KNM comparisons, but it would be better if the separation can be further enlarged, which motivates us to examine different sets of parameters.

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1), span1=0.25", fig.height=3}
# [1] "span1=0.25, span2=0.03, rm_outliers"
# knitr::include_graphics("img/hamby44_v1.png", dpi = 100)
CMPS_hamby44_results$plot[[1]]
```

\autoref{fig:result2_44} shows another Hamby set 44 results, but we changed the value of `npeaks.set` while holding other parameters fixed. In this result, `npeaks.set` is `c(10,6,1)` instead of `c(5,3,1)`. We can see that this results in an increased separation between KM comparisons and KNM comparisons. This shows that the parameters used in the CMPS algorithm could potentially affect the CMPS results, and an open-source implementation of the algorithm will facilitate efforts of cross-validating these parameters.

@@

```{r result2_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(10,6,1), span1=0.25 ", fig.height=3}
# [5] "span1=0.25, span2=0.03, rm_outliers, c(10,6,1)"
# knitr::include_graphics("img/hamby44_c1_v1.png", dpi = 100)
CMPS_hamby44_results$plot[[2]]
```

>>>>

source("../code/func_collection.R")

@@

\wj{Instead of computing the mean CMPS scores of matching and non-matching comparisons, a variation of $\mathrm{\overline{CMPS}_{diff}}$ is to compute the median CMPS scores of matching and non-matching comparisons. Computing the median could be beneficial if extreme values are observed for non-matching comparisons. More examples are provided in the Results section}

@@

```{r hamby252-data-seg, eval=TRUE, echo=FALSE}
# for hamby 252
N <- 8 # the number of different sets of parameters
CMPS_hamby252_results <- list() # a container for everything

#### setup span1
CMPS_hamby252_results$span1 <- as.list(rep(0.25, N))

#### setup signature name
CMPS_hamby252_results$signame <- as.list(rep("sigs25", N))
  # list("sigs75", "sigs25", "sigs25_1062", "sigs15")

#### setup npeaks.set
CMPS_hamby252_results$npeaks.set <- lapply(1:N, function(t) c(5,3,1))

#### setup seg_length
CMPS_hamby252_results$seg_length <- as.list(c(25, 50, 75, 100, 125, 150, 175, 200))

#### setup Tx
CMPS_hamby252_results$Tx <- as.list(rep(25, N))

#### setup outlength
CMPS_hamby252_results$outlength <- vector(mode = "list", length = N)

#### create a plot title and a file name for each set of parameters
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby252_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby252_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252_segment",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}

#### container for the cmps results and corresponding plots
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

#### a global title for Hamby252 plots
com.title252 <- expression(paste(
  "Hamby 252 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

for(i in 1:N){
  CMPS_hamby252_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/hamby252/", 
                        CMPS_hamby252_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

CMPS_hamby252_results_seg <- CMPS_hamby252_results

# generate plots
# for (i in 1:3) {
#   hamby252.cmps <- CMPS_hamby252_results$cmps.table[[i]]
#   
#   hamby252.plot.list <- list()
#   titlee <- CMPS_hamby252_results$titlee[[i]]
#   
#   hamby252.plot.list[[1]] <- hamby252.cmps %>% ggplot() +
#     geom_histogram(aes(x = cmps.max.m,
#                        fill = as.factor(type_truth)), binwidth = 1) +
#     labs(
#       fill = "Comparison Type",
#       x = expression(CMPS[max]),
#       # title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
#       subtitle = titlee
#     ) +
#     scale_x_continuous(breaks = seq(0, 27, 1)) +
#     theme_bw() +
#     theme(panel.grid.minor = element_blank()) +
#     font("x.text", size = 6) +
#     scale_fill_manual(values=c("darkorange", "darkgrey"))
#     
#   hamby252.plot.list[[2]] <- hamby252.cmps %>% ggplot() +
#     geom_histogram(aes(x = cmps.maxbar.m,
#                        fill = as.factor(type_truth)), binwidth = 1) +
#     labs(
#       x = expression(bar(CMPS)[max]),
#       fill = "Comparison Type",
#       # title = expression(paste(
#       #   "Hamby252 - ", bar(CMPS)[max], " Distribution"
#       # )),
#       subtitle = titlee
#     ) +
#     scale_x_continuous(breaks = seq(0, 24, 1)) +
#     theme_bw() +
#     theme(panel.grid.minor = element_blank()) +
#     scale_fill_manual(values=c("darkorange", "darkgrey"))
#   
#   plot <- ggarrange(plotlist = hamby252.plot.list,
#                     nrow = 1,
#                     ncol = 2,
#                     common.legend = TRUE, legend = "bottom")
#   plot <- annotate_figure(plot, 
#                           top = text_grob(com.title252))
#   CMPS_hamby252_results$plot[[i]] <- plot
# }

@@

```{r hamby252-data-npeak, eval=TRUE, echo=FALSE}
N <- 8 # the number of different sets of parameters
CMPS_hamby252_results <- list() # a container for everything

#### setup span1
CMPS_hamby252_results$span1 <- as.list(rep(0.25, N))

#### setup signature name
CMPS_hamby252_results$signame <- as.list(rep("sigs25", N))
# list("sigs75", "sigs25", "sigs25_1062", "sigs15")

#### setup npeaks.set
CMPS_hamby252_results$npeaks.set <- 
  list(c(5),
       c(5,2),
       c(5,3,1),
       c(10,6,2),
       c(6,4,2,1),
       c(10,7,4,2),
       c(10,6,4,2,1),
       c(10,8,6,4,2,1))

#### setup seg_length
CMPS_hamby252_results$seg_length <- 
  # as.list(c(25, 50, 75, 100, 125, 150, 175, 200))
  as.list(rep(50, N))

#### setup Tx
CMPS_hamby252_results$Tx <- as.list(rep(25, N))

#### setup outlength
CMPS_hamby252_results$outlength <- vector(mode = "list", length = N)
CMPS_hamby252_results$outlength[[7]] <- c(60, 90, 120, 150, 180)
CMPS_hamby252_results$outlength[[8]] <- c(50,75,100,125,150,200)

#### create a plot title and a file name for each set of parameters
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby252_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby252_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252_npeak",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}

#### container for the cmps results and corresponding plots
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

#### a global title for Hamby252 plots
com.title252 <- expression(paste(
  "Hamby 252 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

for(i in 1:N){
  CMPS_hamby252_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/hamby252/", 
                        CMPS_hamby252_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

CMPS_hamby252_results_npeak <- CMPS_hamby252_results

```{r hamby44-data-seg, eval=TRUE, echo=FALSE}
N <- 8 # the number of different sets of parameters
CMPS_hamby44_results <- list() # a container for everything

#### setup span1
CMPS_hamby44_results$span1 <- as.list(rep(0.25, N))

#### setup signature name
CMPS_hamby44_results$signame <-
  as.list(rep("sigs25_531", N))

#### setup neapks.set
CMPS_hamby44_results$npeaks.set <-
  lapply(1:N, function(t) c(5,3,1))

#### setup seg_length
CMPS_hamby44_results$seg_length <- as.list(c(30, 45, 61, 90, 122, 150, 180, 210))

#### setup outlength
CMPS_hamby44_results$outlength <- vector(mode = "list", length = N)

#### setup Tx
CMPS_hamby44_results$Tx <- as.list(rep(30, N))

#### create a plot title and a file name for each set of parameters
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby44_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby44_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44_segment",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}

#### container for the cmps results and corresponding plots
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()

#### a global title for Hamby44 plots
com.title44 <- expression(paste(
  "Hamby 44 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

for(i in 1:N){
  CMPS_hamby44_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/hamby44/", 
                        CMPS_hamby44_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

CMPS_hamby44_results_seg <- CMPS_hamby44_results

```{r hamby44-data-npeak, eval=TRUE, echo=FALSE}
N <- 8 # the number of different sets of parameters
CMPS_hamby44_results <- list() # a container for everything

#### setup span1
CMPS_hamby44_results$span1 <- as.list(rep(0.25, N))

#### setup signature name
CMPS_hamby44_results$signame <-
  as.list(rep("sigs25_531", N))

#### setup neapks.set
CMPS_hamby44_results$npeaks.set <-
  list(c(5),
       c(5,2),
       c(5,3,1),
       c(10,6,2),
       c(6,4,2,1),
       c(10,7,4,2),
       c(10,6,4,2,1),
       c(10,8,6,4,2,1))

#### setup seg_length
CMPS_hamby44_results$seg_length <- as.list(rep(61, N))

#### setup outlength
CMPS_hamby44_results$outlength <- vector(mode = "list", length = N)
CMPS_hamby44_results$outlength[[7]] <- c(61, 91, 121, 151, 181)
CMPS_hamby44_results$outlength[[8]] <- c(61, 91, 121, 151, 181, 211)

#### setup Tx
CMPS_hamby44_results$Tx <- as.list(rep(30, N))

#### create a plot title and a file name for each set of parameters
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby44_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby44_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44_npeak",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}

#### container for the cmps results and corresponding plots
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()

#### a global title for Hamby44 plots
com.title44 <- expression(paste(
  "Hamby 44 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

for(i in 1:N){
  CMPS_hamby44_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/hamby44/", 
                        CMPS_hamby44_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

CMPS_hamby44_results_npeak <- CMPS_hamby44_results
```

@@

The sum of squares ratios are 20.64 and 28.96 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. \hh{This indicates, that even though scores from $\mathrm{CMPS_{max}}$ for known-match comparisons are larger than scores from the averaged version of $\mathrm{\overline{CMPS}_{max}}$, these scores achieve a better separation between the two groups of comparisons.}

```{r, echo=FALSE, eval=FALSE}
with(CMPS_hamby252_results_npeak$cmps.table[[3]], {
  tibble(cmps.max, cmps.maxbar) %>% 
    apply(2, compute_var_ratio_anova, label = type_truth, MS = FALSE)
})

with(CMPS_hamby252_results_seg$cmps.table[[2]], {
  tibble(cmps.max, cmps.maxbar, cmps.diff, cmps.diff_scaled) %>% 
    apply(2, compute_var_ratio_anova, label = type_truth, MS = FALSE)
})

```{r plot_computation, echo=FALSE}
# 252 - seg
table.252.cmps_seg <- lapply(CMPS_hamby252_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>% select(cmps.diff:cmps.maxbar_scaled) %>% 
    apply(2, compute_var_ratio_anova, label=type_truth, MS=FALSE)
}) %>% do.call(rbind, .) %>% as.data.frame()

table.252.cmps_seg$seg_length <- unlist(CMPS_hamby252_results_seg$seg_length)

p.seg_252_plot <- table.252.cmps_seg %>%
  pivot_longer(cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
               names_to = "Metric") %>%
  ggplot(aes(x = seg_length, y = value)) +
  geom_line(aes(color = Metric)) + 
  scale_x_continuous(breaks = table.252.cmps_seg$seg_length) +
  theme_bw() +
  ggtitle("A Temp Title") +
  ylab("Sum of Squares Ratio") + 
  xlab("seg_length")

# 252 - npeak
table.252.cmps_npeak <- lapply(CMPS_hamby252_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>% select(cmps.diff:cmps.maxbar_scaled) %>% 
    apply(2, compute_var_ratio_anova, label=type_truth, MS=FALSE)
}) %>% do.call(rbind, .) %>% as.data.frame()

table.252.cmps_npeak$npeak <- CMPS_hamby252_results_npeak$npeaks.set %>% 
  purrr::map(.f = paste, collapse = "-")

p.npeak_252_plot <- table.252.cmps_npeak %>% 
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)])
  ) %>% 
  pivot_longer(cols = c(cmps.diff:cmps.maxbar_scaled),
               names_to = "Metric") %>% 
  ggplot(aes(x=npeak, y=value)) +
  geom_bar(aes(fill = Metric),position = "dodge", stat = "identity") + 
  ylab("Sum of Squares Ratio") +
  coord_flip() + 
  theme_bw()

# 44 - seg
table.44.cmps_seg <- lapply(CMPS_hamby44_results_seg$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>% select(cmps.diff:cmps.maxbar_scaled) %>% 
    apply(2, compute_var_ratio_anova, label=type_truth, MS=FALSE)
}) %>% do.call(rbind, .) %>% as.data.frame()

table.44.cmps_seg$seg_length <- unlist(CMPS_hamby44_results_seg$seg_length)

p.seg_44_plot <- table.44.cmps_seg %>%
  pivot_longer(cols = c(cmps.diff, cmps.diff_scaled, cmps.max, cmps.maxbar),
               names_to = "Metric") %>%
  ggplot(aes(x = seg_length, y = value)) +
  geom_line(aes(color = Metric)) + 
  scale_x_continuous(breaks = table.44.cmps_seg$seg_length) +
  theme_bw() +
  ggtitle("A Temp Title") +
  ylab("Sum of Squares Ratio") + 
  xlab("seg_length")

# 44 - npeak
table.44.cmps_npeak <- lapply(CMPS_hamby44_results_npeak$cmps.table, function(tt) {
  type_truth <- tt$type_truth
  tt %>% select(cmps.diff:cmps.maxbar_scaled) %>% 
    apply(2, compute_var_ratio_anova, label=type_truth, MS=FALSE)
}) %>% do.call(rbind, .) %>% as.data.frame()

table.44.cmps_npeak$npeak <- CMPS_hamby44_results_npeak$npeaks.set %>% 
  purrr::map(.f = paste, collapse = "-")

p.npeak_44_plot <- table.44.cmps_npeak %>% 
  mutate(
    npeak = factor(
      npeak,
      levels = npeak[order(cmps.diff_scaled, decreasing = FALSE)])
  ) %>% 
  pivot_longer(cols = c(cmps.diff:cmps.maxbar_scaled),
               names_to = "Metric") %>% 
  ggplot(aes(x=npeak, y=value)) +
  geom_bar(aes(fill = Metric),position = "dodge", stat = "identity") + 
  ylab("Sum of Squares Ratio") +
  coord_flip() + 
  theme_bw()
```

<!-- this is figure 2 -->
```{r result1_252, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 252; removed outliers; seg\\_length = 50, Tx = 25, npeaks.set = c(5,3,1) ", fig.height=3}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
# knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
# CMPS_hamby252_results$plot[[2]]

p1 <- metric_plot_helper(CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.max", 
                         subtitle = "npeak: 5-3-1, seg_length: 50")
p2 <- metric_plot_helper(CMPS_hamby252_results_seg$cmps.table[[2]], "cmps.maxbar",
                         subtitle = "npeak: 5-3-1, seg_length: 50")
p.result1_252 <- ggarrange(plotlist = list(p1, p2), nrow = 1, ncol = 2,
               common.legend = TRUE, legend = "bottom")
p.result1_252 <- annotate_figure(p.result1_252, top = text_grob(com.title252))
p.result1_252
```

@@

Similar procedures are applied to Hamby set 44, and \autoref{fig:result1_44} shows the distribution of $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. The parameters used in `extract_feature_cmps` are:

@@

Since the resolution of Hamby 44 is set to 1.29 \textmu m per unit, we make `seg_length = 61` and `Tx = 30` to ensure that the setup of Hamby set 44 is similar to that of Hamby set 252, resulting in basis segments of 78.69 \textmu m and the tolerance zone of $\pm 38.7$ \textmu m.

As shown in \autoref{fig:result1_44}, again, we are able to see a clear separation between the known match comparisons and the known non-match comparisons, even though the separation is relatively small compared with that of Hamby set 252, which is also indicated by the sum of squares ratios. For this specific set of parameters, the sum of squares ratios are 8.87 and 10.64 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. This might suggest that we could enlarge the separation in terms of the sum of squares ratio by using other CMPS metrics and other sets of parameters.

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1) ", fig.height=3}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
# knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
# CMPS_hamby252_results$plot[[2]]

p1 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.max",
                         subtitle = "npeak: 5-3-1, seg_length: 61")
p2 <- metric_plot_helper(CMPS_hamby44_results_seg$cmps.table[[3]], "cmps.maxbar",
                         subtitle = "npeak: 5-3-1, seg_length: 61")
p.result1_44 <- ggarrange(plotlist = list(p1, p2), nrow = 1, ncol = 2,
               common.legend = TRUE, legend = "bottom")
p.result1_44 <- annotate_figure(p.result1_44, top = text_grob(com.title44))
p.result1_44
```

**Comparing CMPS metrics and Parameters**

For both Hamby set 252 and 44, we investigated the effects of different sets of parameters. More specifically, we fixed the parameter `npeaks.set` that controls the number of peaks at each segment level and modified the parameter `seg_length` that controls the length of a basis segment. Then we fixed `seg_length` and examined different values of `npeaks.set`. The sum of squares ratios are used to compare among different CMPS metrics and different sets of parameters.

default parameters can be beaten?
cmps.diff, cmps.maxbar and their scaled version are good? we suggest cmps.diff_scaled since it's more comparable and summarizes more information?
Note that the goal of these comparisons is not to determine the optimal set of parameters for the CMPS algorithm, but to show a general procedure that can be used for

```{r}
# CMPS_hamby44_results_seg$cmps.table[[3]] %>% 
#   metric_plot_helper("cmps.max")
# 
# CMPS_hamby252_results_seg$cmps.table[[2]] %>% 
#   metric_plot_helper("cmps.max")
```

```{r seg_252_plot, echo=FALSE, out.width="400px", fig.cap="Some caption", fig.height=3}
p.seg_252_plot
```

```{r npeak_252_plot, echo=FALSE, out.width="400px", fig.cap="Some caption", fig.height=3}
p.npeak_252_plot
```

```{r seg_44_plot, echo=FALSE, out.width="400px", fig.cap="Some caption", fig.height=3}
p.seg_44_plot
```

```{r npeak_44_plot, echo=FALSE, out.width="400px", fig.cap="Some caption", fig.height=3}
p.npeak_44_plot
```

<!-- \autoref{fig:result2_44} shows another Hamby set 44 results, but we changed the value of `npeaks.set` while holding other parameters fixed. In this result, `npeaks.set` is `c(10,6,1)` instead of `c(5,3,1)`. We can see that this results in an increased separation between KM comparisons and KNM comparisons. This shows that the parameters used in the CMPS algorithm could potentially affect the CMPS results, and an open-source implementation of the algorithm will facilitate efforts of cross-validating these parameters. -->

@@

```{r result2_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(10,6,1), span1=0.25 ", fig.height=3}
# [5] "span1=0.25, span2=0.03, rm_outliers, c(10,6,1)"
# knitr::include_graphics("img/hamby44_c1_v1.png", dpi = 100)
# CMPS_hamby44_results$plot[[2]]
```
